Road traffic sound level estimation from realistic urban sound mixtures by Non-negative Matrix Factorization

### 2018-03-05 paper/gloaguenAppliedNMF.tex
minor updates

The urban sound scenes are taken from 76 recordings from 2 to 5 min, achieved in the \nth{13} district of Paris (France) at 19 different locations \footnote{Recordings were made as part of the Grafic project funded by Ademe}, which cover four various sound environments (Figure \ref{fig:map_grafic}), see Figure \ref{fig:map_grafic}. A complete description of the experimental protocol can be found in \cite{aumond_modelling_2017}. Two of the 76 recordings are rejected for the analysis because the audio files were corrupted, resulting in 74 valid audio files assumed as representative of the variety of sound environments. The recordings are listened and categorized within four different sound environments, as proposed in \cite{can_describing_2015}: park (P, 8 audio files), quiet street (Q, 35 audio files), noisy street (N, 23 audio files) and very noisy street (vN, 8 audio files). Then, each audio file is annotated, noting the start and end time of each sound event along with its sound class. The aim of the annotation phase is next to transcribe the recordings, in order to obtain simulated sound scenes with the same distribution of sound events as the recordings and therefore as close as possible to the realistic scenes.

@@

First the spectrogram of each audio file is computed (window $w = 2^{12}$ with 50 $\%$ overlap). The spectrogram is then cut in multiple temporal frames of $w_t  = \lbrace$0.5, 1$\rbrace$ second duration. In each cut spectrogram, the root mean square on each frequency bin is calculated to obtained a spectrum of $F \times 1$ dimension. This method allows the description of the audio sample with finer spectra and then having the different characteristic pitches of the traffic spectra. An illustrative example on a 3 seconds sample is displayed in Figure \ref{fig:example_dictionary}. From the 53 audio files, we obtain respectively for $w_t\in$ $\lbrace$0.5, 1$\rbrace$ second, 2218 and 1109 elements. A $K$-mean clustering algorithm is applied to reduce these dimensions to $K \in \lbrace$25, 50, 100, 200$\rbrace$ in order to avoid redundant information and decrease the computation time. The obtained $K$ clusters compose then the elements of $\mathbf{W}$. Furthermore, the case where each audio generates one spectrum from its spectrogram is added ($w_t = all$). Here, by the added approach, the dictionary elements are based on the spectral envelops of the audio samples. For this case, having 53 audio samples, the number of elements in $\mathbf{W}$ with the $K$-mean clustering algorithm is reduced to $K \in \lbrace$25, 50$\rbrace$.

The obtained dictionary is expressed in third octave bands to decrease the dimensionality without deteriorate the spectral content. A previous experimental validation revealed that the use of third octave bands, instead of linear spectra, does not impact the performances of the chosen estimator in this study.  Finally, each basis of $\mathbf{W}$ is normalized as $ \Vert \mathbf{w}_k \Vert = 1$ where $\Vert \bullet \Vert$ is the $\ell_1$ norm. Table \ref{tab:experimental_factorsNMF} summarizes the different modalities of the two experimental factors ($K$ and $w_t$).

@@

NMF is performed for 3 $\beta$-divergences: $\beta$ = 2 (euclidean distance), $\beta$ = 1 (Kullback-Leibler divergence) and $\beta = 0$ (Itakura-Sa\"ito divergence). The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are displayed in a logarithmic scale through a third band octave representation that reduces the high frequency predominance where the traffic component is absent. In addition, as the number of frequency bins is reduced ($F$ = 29), the computation time is reduced too. 400 iterations are performed to get a stabilized results. For SEM-NMF, the number of elements in $\mathbf{W_r}$ is fixed at $J = 2$. For hard thresholding, the threshold value, $t_h$ is defined between 0.30 and 0.60 with a 0.01 step. 
Each unique association of modalities between each experimental factor forms a setting. For the filter estimator, 24 settings are computed (4 $\times$ 6). For SUP and SEM-NMF, 240 settings are computed (4 $\times$ 2 $\times$ (2 $\times$ 4 + 1 $\times$ 2) $\times$ 3). Finally, for TI-NMF, the number of settings is much higher (3720) due to the multiple possible combinations between the threshold values (4 $\times$ 1 $\times$ (2 $\times$ 4 + 1 $\times$ 2) $\times$ 3 $\times$ 31).
The summarize of the experimental factors and their different modalities is displayed on Tables \ref{tab:experimental_factorsFilter} and \ref{tab:experimental_factorsNMF}.

@@

where the other experimental factors (method, $f_c$, $K$, $w_t$, $\beta$, threshold value $t_h$) are fixed.

@@

\begin{table*}[t]
\centering
\caption{Best $mMAE$ errors according to the experimental factors $\beta$ and \textit{method} (in bold letter, the lowest error).}
\begin{tabular}{@{}ccccccc@{}}
\toprule
\textbf{method} & $f_c$ (kHz) & $\mathbf{\beta}$ & $\mathbf{K}$ & $\mathbf{w_t}$ (s) &  $\mathbf{t_h}$ &  \textbf{$mMAE$} (dB) \\ \midrule
filter & 20 & - & - & - & - & 3.76 ($\pm$ 4.35) \\
filter & 0.5 & - & - & - & - & 2.14 ($\pm$ 1.83) \\ 
\hline \hline
SUP-NMF & - & 0 & 200 & 0.5 & - & 4.06 ($\pm$ 4.69) \\
SUP-NMF & - & 1 & 200 & 0.5 & - & 2.79 ($\pm$ 3.38) \\
SUP-NMF & - & 2 & 25 & 1 & - & 2.32  ($\pm$ 2.80) \\ 
\hline \hline
SEM-NMF & - & 0 & 200 & 1 & - & 2.05 ($\pm$ 0.70) \\
SEM-NMF & - & 1 & 200 & 1 & - & 1.94 ($\pm$ 0.38) \\
SEM-NMF & - & 2 & 200 & 1 & - & 2.39 ($\pm$ 1.23) \\ 
\hline \hline
TI-NMF & - & 0 & 50 & all & 0.39 & 1.44 ($\pm$ 1.09)\\
TI-NMF & - & 1 & 200 & 0.5 & 0.35 & 1.40 ($\pm$ 1.22)\\
\textbf{TI-NMF }& - & \textbf{2} & \textbf{200} & \textbf{0.5} & \textbf{0.33} & \textbf{1.32 ($\pm$ 1.25)}\\
\bottomrule
\end{tabular}
\label{tab:results}
\end{table*}

@@

TI-NMF is the approach with the lowest global error ($<$ 1.50 dB). The best result is obtained for TI-NMF ($MAE$ = 1.31 $\pm$ 1.26 dB) with $\beta = 2$, $K$ = 200, $w_t$ = 0.5 s and as threshold value $t_h$ = 0.33. This combination of settings offers the most fitted method to be adapted to all the sound environments. 
Furthermore, on the dictionary creation, except SEM-NMF where all the best methods according to $\beta$ use the same dictionary, no specific dictionary form through all the method, is used. SUP and SEM-NMF privilege a high number of element ($K$ = 200) with a fine description of the audio samples ($w_t \in \lbrace 0.5$,$1 \rbrace$ second). For TI-NMF, the composition of the initial dictionary is more diverse both on the $K$ value and on the finesse  of the description ($w_t \in \lbrace all$,$0.5$, $1 \rbrace$ second).

@@

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{figures/mea_grafic_bar.pdf}
\caption{$MAE$ errors according each sound environment for the best combination of the LP filter ($f_c$ = 500 Hz), SUP-NMF ($\beta$ = 2, $K$ = 25, $w_t$ = 1 second), SEM-NMF ($\beta$ = 1, $K$ = 200, $w_t$ = 1 second) and TI-NMF ($\beta$ = 2, $K$ = 200, $w_t$ = 0.5 second, $t_{h} = 0.33$)}
\label{fig:mae_env}
\end{figure}

@@

Besides having the lowest $mMAE$ error, TI-NMF presents the most performing results. The park environment is the only case where an other NMF approach out-performed significantly TI-NMF ($MAE$ = 3.05 dB). In this sound environment, the traffic dictionary is then composed, in average, of more than half the 200 elements in $\mathbf{W'}$ (136 elements). For the rest of the sound environments, TI-NMF has the lowest error. For very noisy environment the error is even very low ($MAE=$ 0.31 dB). In this case, in average, $\mathbf{W'}$ is composed of 198 traffic elements. This method out-performed SUP-NMF since, as $\mathbf{W_0}$ is updated, the final dictionary $\mathbf{W'}$ is then most suited to the sound scene than a fixed dictionary. The advantage to have an unique dictionary suited to each sound scene makes TI-NMF very performing when traffic is predominant while, by the thresholding, it limits the dictionary derivation when the traffic is less present.

@@

The results confirm the interest of this method on such sound environments. It easily takes into account the overlap between the multiple sound sources present in cities and is suited to monophonic sensor networks. Different NMF formula have been studied through the supervised and semi-supervised approach. On all the sound environments, these common approaches reveal to be not sufficiently efficient: supervised NMF approach, with its fixed dictionary, does not succeed to estimate correctly the traffic sound level especially when this sound source is quiet, while semi-supervised approach with the presence of a mobile part in the dictionary is the best estimator for \textit{park} environments but failed on heavily traffic scenes. Finally, the proposed approach, namely Thresholded Initialized NMF, achieved the lowest error with an initialized updated dictionary. By considering the elements the most similar to traffic spectra with a hard thresholding, it makes it possible to adapt specifically the dictionary to the sound scenes. Consequently, in the case where the emplacement or the sound environment of the sensors cannot be identified (for instance within a mobile measurement framework), TI-NMF, with $\beta = 2$, $K$ = 200, $w_t = 1$ second, $t_{h} = 0.33$, seems the best generic framework, for the sound corpus studied. But if the sound environment can be identified through a prior analysis, or based on positioning data \cite{can2015noise,lavandier2016urban}, it seems possible to adapt the proposed work by selecting the most efficient approach (for instance SEM-NMF within parks), in order to further reduce the error in the estimated road traffic sound levels.
Further analyses are required to extend the proposed method to other sound sources, such as birds or voices sounds, by replacing or adding elements in the built dictionary. This would prove useful in the context of muti-source noise mapping that is  gaining interest \cite{aumond2017Probabilistic, aletta2015soundscape}. Finally, the selected parameters stand for the corpus of sound mixtures of this study. Further analyses on various corpus of sound scenes are needed to test the robustness of the method, and select the most relevant approaches for specific sound environments (predominance of water or industrial sounds, rural environments \dots).

>>>>

The urban sound scenes are taken from 76 recordings from 2 to 5 min, achieved in the \nth{13} district of Paris (France) at 19 different locations \footnote{Recordings were made as part of the Grafic project funded by Ademe}, which cover four various sound environments (Figure \ref{fig:map_grafic}), see Figure \ref{fig:map_grafic}. A complete description of the experimental protocol can be found in \cite{aumond_modelling_2017}. Two of the 76 recordings are rejected for the analysis because the audio files were corrupted, resulting in 74 valid audio files assumed as representative of the variety of sound environments. The recordings are listened and categorized within four different sound environments, as proposed in \cite{can_describing_2015}: park (P, 8 audio files), quiet street (Q, 33 audio files), noisy street (N, 24 audio files) and very noisy street (vN, 9 audio files). Then, each audio file is annotated, noting the start and end time of each sound event along with its sound class. The aim of the annotation phase is next to transcribe the recordings, in order to obtain simulated sound scenes with the same distribution of sound events as the recordings and therefore as close as possible to the realistic scenes.

@@

First the spectrogram of each audio file is computed (window $w = 2^{12}$ with 50 $\%$ overlap). The spectrogram is then cut in multiple temporal frames of $w_t  = \lbrace$0.5, 1$\rbrace$ second duration. In each cut spectrogram, the root mean square on each frequency bin is calculated to obtained a spectrum of $F \times 1$ dimension. This method allows the description of the audio sample with finer spectra and then having the different characteristic pitches of the traffic spectra. An illustrative example on a 3 seconds sample is displayed in Figure \ref{fig:example_dictionary}. From the 53 audio files, we obtain respectively for $\mathbf{w_t}\in$ $\lbrace$0.5, 1$\rbrace$ second, 2218 and 1109 elements. A $\mathbf{K}$-mean clustering algorithm is applied to reduce these dimensions to $\mathbf{K} \in \lbrace$25, 50, 100, 200$\rbrace$ in order to avoid redundant information and decrease the computation time. The obtained $\mathbf{K}$ clusters compose then the elements of $\mathbf{W}$. Furthermore, the case where each audio generates one spectrum from its spectrogram is added ($\mathbf{w_t} = all$). Here, by the added approach, the dictionary elements are based on the spectral envelops of the audio samples. For this case, having 53 audio samples, the number of elements in $\mathbf{W}$ with the $\mathbf{K}$-mean clustering algorithm is reduced to $\mathbf{K} \in \lbrace$25, 50$\rbrace$.

The obtained dictionary is expressed in third octave bands to decrease the dimensionality without deteriorate the spectral content. A previous experimental validation revealed that the use of third octave bands, instead of linear spectra, does not impact the performances of the chosen estimator in this study.  Finally, each basis of $\mathbf{W}$ is normalized as $ \Vert \mathbf{w}_k \Vert = 1$ where $\Vert \bullet \Vert$ is the $\ell_1$ norm. Table \ref{tab:experimental_factorsNMF} summarizes the different modalities of the two experimental factors ($\mathbf{K}$ and $\mathbf{w_t}$).

@@

NMF is performed for 3 $\beta$-divergences: $\beta$ = 2 (euclidean distance), $\beta$ = 1 (Kullback-Leibler divergence) and $\beta = 0$ (Itakura-Sa\"ito divergence). The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are displayed in a logarithmic scale through a third band octave representation that reduces the high frequency predominance where the traffic component is absent. In addition, as the number of frequency bins is reduced ($F$ = 29), the computation time is reduced too. 400 iterations are performed to get a stabilized results. For SEM-NMF, the number of elements in $\mathbf{W_r}$ is fixed at $J = 2$. For hard thresholding, the threshold value, $\mathbf{t_h}$ is defined between 0.30 and 0.60 with a 0.01 step. 
Each unique association of modalities between each experimental factor forms a setting. For the filter estimator, 24 settings are computed (4 $\times$ 6). For SUP and SEM-NMF, 240 settings are computed (4 $\times$ 2 $\times$ (2 $\times$ 4 + 1 $\times$ 2) $\times$ 3). Finally, for TI-NMF, the number of settings is much higher (3720) due to the multiple possible combinations between the threshold values (4 $\times$ 1 $\times$ (2 $\times$ 4 + 1 $\times$ 2) $\times$ 3 $\times$ 31).
The summarize of the experimental factors and their different modalities is displayed on Tables \ref{tab:experimental_factorsFilter} and \ref{tab:experimental_factorsNMF}.

@@

where the other experimental factors (method, $f_c$, $\mathbf{K}$, $\mathbf{w_t}$, $\beta$, threshold value $\mathbf{t_h}$) are fixed.

@@

\begin{table*}[t]
\centering
\caption{Best $mMAE$ errors according to the experimental factors $\beta$ and \textit{method} (in bold letter, the lowest error).}
\begin{tabular}{@{}ccccccc@{}}
\toprule
\textbf{method} & $f_c$ (kHz) & $\mathbf{\beta}$ & $\mathbf{K}$ & $\mathbf{w_t}$ (s) &  $\mathbf{t_h}$ &  \textbf{$mMAE$} (dB) \\ \midrule
filter & 20 & - & - & - & - & 3.76 ($\pm$ 4.35) \\
filter & 0.5 & - & - & - & - & 2.14 ($\pm$ 1.83) \\ 
\hline \hline
SUP-NMF & - & 0 & 200 & 0.5 & - & 4.06 ($\pm$ 4.69) \\
SUP-NMF & - & 1 & 200 & 0.5 & - & 2.79 ($\pm$ 3.38) \\
SUP-NMF & - & 2 & 25 & 1 & - & 2.32  ($\pm$ 2.80) \\ 
\hline \hline
SEM-NMF & - & 0 & 200 & 1 & - & 2.05 ($\pm$ 0.70) \\
SEM-NMF & - & 1 & 200 & 1 & - & 1.94 ($\pm$ 0.38) \\
SEM-NMF & - & 2 & 200 & 1 & - & 2.39 ($\pm$ 1.23) \\ 
\hline \hline
TI-NMF & - & 0 & 25 & 1 & 0.39 & 1.42 ($\pm$ 0.89)\\
TI-NMF & - & 1 & 100 & 1 & 0.35 & 1.38 ($\pm$ 0.88)\\
\textbf{TI-NMF }& - & \textbf{2} & \textbf{200} & \textbf{0.5} & \textbf{0.32} & \textbf{1.24 ($\pm$ 1.24)}\\
\bottomrule
\end{tabular}
\label{tab:results}
\end{table*}

@@

TI-NMF is the approach with the lowest global error ($<$ 1.50 dB). The best result is obtained for TI-NMF ($MAE$ = 1.24 ($\pm$ 1.24) dB) with $\beta = 2$, $\mathbf{K}$ = 200, $\mathbf{w_t}$ = 0.5 s and as threshold value $\mathbf{t_h}$ = 0.32. This combination of settings offers the most fitted method to be adapted to all the sound environments. 
Furthermore, on the dictionary creation, except SEM-NMF where all the best methods according to $\beta$ use the same dictionary, no specific dictionary form through all the method, is used. SUP and SEM-NMF privilege a high number of element ($\mathbf{K}$ = 200) with a fine description of the audio samples ($\mathbf{w_t} \in \lbrace 0.5$,$1 \rbrace$ second). For TI-NMF, the composition of the initial dictionary is more diverse both on the $\mathbf{K}$ value and on the finesse  of the description ($\mathbf{w_t} \in \lbrace all$,$0.5 \rbrace$ second).

@@

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{figures/mea_grafic_bar.pdf}
\caption{$MAE$ errors according each sound environment for the best combination of the LP filter ($f_c$ = 500 Hz), SUP-NMF ($\beta$ = 2, $\mathbf{K}$ = 25, $\mathbf{w_t}$ = 1 second), SEM-NMF ($\beta$ = 1, $\mathbf{K}$ = 200, $\mathbf{w_t}$ = 1 second) and TI-NMF ($\beta$ = 2, $\mathbf{K}$ = 200, $\mathbf{w_t}$ = 0.5 second, $\mathbf{t_{h}} = 0.32$)}
\label{fig:mae_env}
\end{figure}

@@

Besides having the lowest $mMAE$ error, TI-NMF presents the most performing results. The park environment is the only case where an other NMF approach out-performed significantly TI-NMF ($MAE$ = 2.95 dB). In this sound environment, the traffic dictionary is then composed, in average, of more than half the 200 elements in $\mathbf{W'}$ (136 elements). For the rest of the sound environments, TI-NMF has the lowest error. For very noisy environment the error is even very low ($MAE=$ 0.28 dB). In this case, in average, $\mathbf{W'}$ is composed of 198 traffic elements. This method out-performed SUP-NMF since, as $\mathbf{W_0}$ is updated, the final dictionary $\mathbf{W'}$ is then most suited to the sound scene than a fixed dictionary. The advantage to have an unique dictionary suited to each sound scene makes TI-NMF very performing when traffic is predominant while, by the thresholding, it limits the dictionary derivation when the traffic is less present.

@@

The results confirm the interest of this method on such sound environments. It easily takes into account the overlap between the multiple sound sources present in cities and is suited to monophonic sensor networks. Different NMF formula have been studied through the supervised and semi-supervised approach. On all the sound environments, these common approaches reveal to be not sufficiently efficient: supervised NMF approach, with its fixed dictionary, does not succeed to estimate correctly the traffic sound level especially when this sound source is quiet, while semi-supervised approach with the presence of a mobile part in the dictionary is the best estimator for \textit{park} environments but failed on heavily traffic scenes. Finally, the proposed approach, namely Thresholded Initialized NMF, achieved the lowest error with an initialized updated dictionary. By considering the elements the most similar to traffic spectra with a hard thresholding, it makes it possible to adapt specifically the dictionary to the sound scenes. Consequently, in the case where the emplacement or the sound environment of the sensors cannot be identified (for instance within a mobile measurement framework), TI-NMF, with $\beta = 2$, $\mathbf{K}$ = 200, $\mathbf{w_t} = 0.5$ second, $\mathbf{t_{h}} = 0.32$, seems the best generic framework, for the sound corpus studied. But if the sound environment can be identified through a prior analysis, or based on positioning data \cite{can2015noise,lavandier2016urban}, it seems possible to adapt the proposed work by selecting the most efficient approach (for instance SEM-NMF within parks), in order to further reduce the error in the estimated road traffic sound levels.
Further analyses are required to extend the proposed method to other sound sources, such as birds or voices sounds, by replacing or adding elements in the built dictionary. This would prove useful in the context of muti-source noise mapping that is  gaining interest \cite{aumond2017Probabilistic, aletta2015soundscape}. Finally, the selected parameters stand for the corpus of sound mixtures of this study. Further analyses on various corpus of sound scenes are needed to test the robustness of the method, and select the most relevant approaches for specific sound environments (predominance of water or industrial sounds, rural environments \dots).
